Reconstructing Propositional Proofs in Type Theory

### 2017-11-20 paper/sections/abstract.tex
[ paper ] lots of changes.

We describe a syntactical proof-reconstruction approach to verify proofs
generated by \Metis prover to problems in classical propositional logic. To
verify \Metis proofs, we developed a tool and two Agda libraries able to
translate such derivations to \Agda proof-terms. We formalise in type theory
each inference rule found in the \Metis derivations. This allowed us to type-
check with \Agda, \Metis proofs step-by-step.

% We leave it for future work, the opportunity to
% extend this work to support other provers, proof-assistants, and formalisms.

>>>>

We describe a syntactical proof-reconstruction approach to verify derivations
generated by \Metis prover to theorems in classical propositional logic. To
verify such derivations, we formalise in type theory each inference rule
of the \Metis reasoning. We developed a tool jointly with two \Agda libraries
to translate \Metis derivations to \Agda proof-terms.
These developments allowed us to type-check with \Agda, \Metis proofs
step-by-step.

### 2017-11-20 paper/sections/introduction.tex
This closes #40
[ paper ] lots of changes.

An automatic theorem prover (henceforth \ATP) is a program that
intends to prove conjectures from axioms and inference rules of some
logical system. In the last decades, \ATPs are fast becoming a key
instrument in different disciplines and real applications (\eg,
verifying railway interlocking systems~\cite{Kanso2012}). Since some
bugs have been found in these programs (see, for example,
\cite{Keller2013,Bohme2011,Fleury2014}), researchers and users from
academy and industry have shown an increased interest to
\emph{formally} prove the validity of \ATPs' results.

@@

Proof-reconstruction address this problem. It becomes in mostly cases
a reverse engineering task to verify the prover reasoning. That
usually happen when the reconstruction is made by another and not by
the developers of the \ATP. The presentation of the proofs generated
by the prover play an important role in proof-reconstruction. To
verify such a proof, it is convenient to have it in a consistent
format, that is, a full script describing step-by-step with exhaustive
details and without ambiguities, the derivation to get the proof. For
classical propositional logic (henceforth \CPL) from a list of at
least forty \ATPs---available from the Web service \name{SystemOnTPTP}
of the TPTP
World\footnote{\url{http://www.cs.miami.edu/~tptp/}.}---just few of
them show their proofs.

One approach to address the proof-reconstruction problem is
integrating the \ATP, the \emph{source} system, with a
proof-assistant, the \emph{target} system. The target system is the
proof \emph{checker} in charge to verify the source system reasoning.
These proof-assistants allow us to formalize the logical system used
in the proofs, \ie, logical constants, axioms, inference rules,
hypotheses, and theorems. A proof-reconstruction tool provides such an
integration, translating the derivation generated by the prover into
the formalism of the proof-assistant.

Previous studies have reported proof-reconstruction using
proof-assistants based on higher-order logic where the development is
at a mature
stage~\cite{paulson2010three,hurlin07practical,kaliszyk2013}. Another
approaches has been proposed for proof-reconstructing based on type
theory in~\cite{Bezem2002,kanso2016light,Keller2013}.

@@

% Finally, in appendix \textsc{A}, we provide a small example of the
% proof-reconstruction workflow.

The source code accompanying this paper (programs, libraries, and
examples) is available at \verb!GitHub!:
\begin{itemize}
  \item The \Athena program that translates proofs generated by \Metis to \Agda:
  \url{http://github.com/jonaprieto/athena}.
  \item The \verb!agda-prop! library as a formalization in \Agda for
    classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.
\item The \verb!agda-metis! library as a formalization in \Agda to
  justify \Metis proofs of classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.
\end{itemize}

We carried out proof-reconstruction for \Metis~2.3
(release~20170822). The \Athena program was tested with \prg{GHC}~8.2.1.
The \verb!agda-prop! and \verb!agda-metis! libraries were tested with
\Agda~2.5.3 and \Agda standard library~0.14.

>>>>

An automatic theorem prover (henceforth \ATP) is a program that
intends to prove conjectures from axioms and inference rules of some
logical system. In the last decades, \ATPs are fast becoming a key
instrument in different disciplines and real applications (\eg,
verifying a railway interlocking system, an operating-system kernel,
or a pseudo-random number generator for cryptography). Since some
programming errors have been found in these programs (see, for
example, \cite{Keller2013,Bohme2011,Fleury2014}), researchers and
users from academy and industry have shown an increased interest to
\emph{formally} prove the validity of \ATPs' results.

@@

Proof-reconstruction addresses this problem. It becomes in mostly
cases a reverse engineering task to verify the prover reasoning. That
usually happens when the reconstruction is made by another and not by
the developers of the \ATP. The presentation of the proofs generated
by the prover plays an important role in proof-reconstruction. To
verify such a proof, it is convenient to have it in a consistent
format, that is, a full script describing the derivation step-by-step
with  exhaustive details and without ambiguities. For example, for
classical propositional logic (henceforth \CPL) from a list of at
least forty \ATPs---available from the Web service
\name{SystemOnTPTP} of the TPTP World\footnote{\url{http://www.cs.miami.edu/~tptp/}.}---just few of
them show their proofs.

One approach to address the proof-reconstruction problem is
justifying each inference rule of the \emph{source} system, the \ATP,
with a proof-assistant, the \emph{target} system. The target system
is the proof \emph{checker} in charge to verify the source system
reasoning.
These proof-assistants allow us to formalize the logical system used
in the proofs, \ie, logical constants, axioms, inference rules,
hypotheses, and theorems. A proof-reconstruction tool provides such
an integration, translating the derivation generated by the prover
into the formalism of the proof-assistant.

Previous studies have reported proof-reconstruction using proof-
assistants based on higher-order logic where the development is at a
mature stage~\cite{paulson2010three,hurlin07practical,kaliszyk2013}.
Another approaches has been proposed for proof-reconstructing based
on type theory in~\cite{Bezem2002,kanso2016light,Keller2013}.

@@

The source code accompanying this paper (programs, libraries, and
examples) is available at \verb!GitHub!:

\begin{itemize}
  \item The \Athena program that translates proofs generated by \Metis
  to \Agda code: \url{http://github.com/jonaprieto/athena}.

  \item The \verb!agda-prop! library as a formalization in \Agda for
    classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.

  \item The \verb!agda-metis! library as a formalization in \Agda to
  justify \Metis proofs of classical propositional logic:
  \url{http://github.com/jonaprieto/agda-prop}.

\end{itemize}

The proof-reconstruction tool \Athena was tested with \prg{GHC}~8.2.1.
Both libraries, \verb!agda-prop! and \verb!agda-metis! were tested with
\Agda~2.5.3 and \Agda standard library~0.14.
\Athena jointly with \verb!agda-prop! and \verb!agda-metis! are able to reconstruct propositional proofs of \Metis~2.3 (release~20171021).

### 2017-11-20 paper/sections/metis-language-and-proofs.tex
[ paper ] fixed typos in metis descrp..
[ paper ] lots of changes.

\begin{figure}
\begin{equation*}
  \begin{bprooftree}
    \AxiomC{}
    \RightLabel{axiom}
    \UnaryInfC{$Γ ⊢ φ$}
  \end{bprooftree}
  \qquad
  \begin{bprooftree}
    \AxiomC{}
    \RightLabel{assume $φ$}
    \UnaryInfC{$Γ ⊢ φ ∨ ¬ φ$}
  \end{bprooftree}
  \end{equation*}
  \vskip2mm
  \begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ l ∨ φ$}
    \AxiomC{$Γ ⊢ ¬ l ∨ ψ$}
    \RightLabel{resolve $l$}
    \BinaryInfC{$Γ ⊢ φ ∨ ψ$}
  \end{bprooftree}
\end{equation*}
\caption{Propositional logic inference rules of the \texttt{Metis} prover.}
\label{fig:metis-inferences}
\end{figure}

@@

The \TPTP language is the input language to encode problems used by \Metis.
It includes the first-order form (denoted by \name{fof}) and clause normal form (denoted by \name{cnf}) formats~\cite{sutcliffe2009}.
The \TPTP syntax describes a well-defined grammar to handle annotated
formulas with the following form:

@@

where the \name{language} can be \name{fof} or \name{cnf}. The
\name{name} serves to identify the formula within the problem. Each
formula assumes a \name{role}, this could be an \name{axiom},
\name{conjecture}, \name{definition}, \name{plain} or an
\name{hypothesis}.  The formulas include the constants
\verb!$true! and \verb!$false!, the negation unary operator~(\verb!~!),
and the binary connectives \verb!&!, \verb!|!, \verb!=>!, and
(\verb!<=>!) to represent ($⊤$, $⊥$, $¬$, $∧$, $∨$, $⇒$), and ($⇔$)
respectively.

@@

\begin{verbatim}
  fof(premise, axiom, p).
  fof(goal, conjecture, p).
  fof(sg0, p, inf(strip, [goal])).
  fof(neg0, ¬ p, inf(negate, [sg0])).
  fof(norm0, ¬ p, inf(canonicalize, [neg0])).
  fof(norm1, p, inf(canonicalize, [premise])).
  fof(norm2, ⊥, inf(simplify, [norm0, norm1]))
  cnf(refute0, ⊥, inf(canonicalize, [norm2])).
\end{verbatim}

@@

A proof generated by \Metis encodes a natural
deduction proof.
With the inference rules in Fig.~\ref{tab:agda-metis-table}
as the only valid deduction steps, \Metis
attempts to prove conjectures by refutation (\ie,
\emph{falsium} in the root of the \TSTP derivation).

>>>>

\begin{figure}
\begin{equation*}
  \begin{bprooftree}
    \AxiomC{}
    \RightLabel{axiom}
    \UnaryInfC{$Γ ⊢ φ$}
  \end{bprooftree}
  \qquad
  \begin{bprooftree}
    \AxiomC{}
    \RightLabel{assume $φ$}
    \UnaryInfC{$Γ ⊢ φ ∨ ¬ φ$}
  \end{bprooftree}
  \end{equation*}
  \vskip2mm
  \begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ ℓ ∨ φ$}
    \AxiomC{$Γ ⊢ ¬ ℓ ∨ ψ$}
    \RightLabel{resolve $ℓ$}
    \BinaryInfC{$Γ ⊢ φ ∨ ψ$}
  \end{bprooftree}
\end{equation*}
\caption{Propositional logic inference rules of the \texttt{Metis} prover.}
\label{fig:metis-inferences}
\end{figure}

@@

The \TPTP language is the input language to encode problems used by \Metis.
It includes the first-order form (denoted by \name{fof}) and clause normal form (denoted by \name{cnf}) formats~\cite{sutcliffe2009}.
The \TPTP syntax\footnote{See the complete grammar
at \url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}} describes a well-defined grammar to handle annotated
formulas with the following form:

@@

where the \name{language} can be \name{fof} or \name{cnf}. The
\name{name} serves to identify the formula within the problem. Each
formula assumes a \name{role}, this could be an \name{axiom},
\name{conjecture}, \name{definition}, \name{plain} or an
\name{hypothesis}.

The formulas include the constants
\verb!$true! and \verb!$false!, the negation unary
operator~(\verb!~!), and the binary connectives
(\verb!&!, \verb!|!, \verb!=>!) to represent
($⊤$, $⊥$, $¬$, $∧$, $∨$, $⇒$) respectively.

@@

\begin{verbatim}
  fof(premise, axiom, p).
  fof(goal, conjecture, p).
  fof(s₀, p, inf(strip, goal)).
  fof(neg₀, ¬ p, inf(negate, s₀)).
  fof(n₀, ¬ p, inf(canonicalize, neg₀)).
  fof(n₁, p, inf(canonicalize, premise)).
  fof(n₂, ⊥, inf(simplify, [n₀, n₁]))
  cnf(r₀, ⊥, inf(canonicalize, n₂)).
\end{verbatim}

@@

A proof generated by \Metis encodes a natural
deduction proof, Fig.~\ref{fig:metis-example} is an example of such kind of proof. With the inference rules in
Fig.~\ref{tab:agda-metis-table}
as the only valid deduction steps, \Metis
attempts to prove conjectures by refutation (\ie,
\emph{falsium} in the root of the \TSTP derivation).

### 2017-11-20 paper/sections/proof-reconstruction.tex
Updating the paper in master for the visibility. (#42)
* [ paper ] Eq removed.

* resolve fixing

* [ paper ] last changes.

* [ paper ] lots of changes.

* ruu

* ppp

* [ paper ] lots of changes.

* [ paper ] last changes.

* [ paper ] some messages added.

* [ paper ] fixed subgoal description.

* [ paper ] fixed rm funnctions and example added.

* [ paper ] fixed normalized formula description.

* [ paper ] canon-and fixed.

* [ paper ] fixed form function text description.

* [ paper ] fixed canoncialize theorem.

* [ paper ] fixed conjunct example.

* [ paper ] fixed minor typos.

* [ pa

\begin{myexamplenum}
\label{ex:inference-rule-pattern}
\Metis has a inference rule named \conjunct. We reconstruct this
inference rule as follow. We define in \Agda the function
\fconjunct which has the following pattern:
\begin{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fconjunct : \Source → \Target → \Prop\\
  &\begin{array}{ll}
  \fconjunct~φ~ψ\ &=
      \begin{cases}
      ψ, &\text{if }\fconjunct \text{ built }ψ\text{ by applying valid}\\
         &\text{transformations or other rules to }φ;\\
      φ, &\text{otherwise;}
      \end{cases}
  \end{array}
  \end{aligned}
\end{equation*}

To justify all transformations done by the \fconjunct rule, we prove
its soundness with a theorem. This theorem would be
\begin{equation*}
  \text{If }Γ ⊢ φ \text{ then }Γ ⊢ \fconjunct~φ~ψ, \text{ where } ψ : \Target.
\end{equation*}
\end{myexamplenum}

>>>>

\begin{myexamplenum}
\label{ex:inference-rule-pattern}
Let \texttt{metisRule} be a \Metis inference rule. To reconstruct this rule,
we define the function \fun{metisRule} in type theory that follows the pattern:

\begin{equation*}
  \begin{aligned}
  &\hspace{.495mm}\fun{metisRule} : \Source → \Target → \Prop\\
  &\begin{array}{ll}
  \fun{metisRule}~φ~ψ\ &=
      \begin{cases}
      ψ, &\text{if the conclusion}ψ\text{ can be}\text{ derived by applying certain inference}\\
         &\text{rules to the premise }φ;\\
      φ, &\text{otherwise;}
      \end{cases}
  \end{array}
  \end{aligned}
\end{equation*}

To justify all transformations done by the \fun{metisRule} rule, we prove
its soundness with a theorem like the following:

\begin{equation*}
  \text{If }Γ ⊢ φ \text{ then }Γ ⊢ \fun{metisRule}~φ~ψ, \text{ where } ψ : \Target.
\end{equation*}

\end{myexamplenum}

### 2017-11-20 paper/sections/proof-reconstruction/clausify.tex
[ paper ] some corrections added.
[ paper ] lots of changes.

\subsubsection{Clausification.}
\label{sssec:clausification}

The \clausify rule transforms a formula into its clausal normal form.
It is important to notice that this kind of conversion between one formula
to its clausal normal form is not unique, and \Metis has a customized
approach to perform that transformation. Hence, we perform a
reordering of the conjunctive normal form for the source formula.

\begin{myexamplenum}
In the following \TSTP derivation by \Metis, we see how
\clausify transforms the \texttt{norm0} formula to get \texttt{norm1} formula.

\begin{verbatim}
  fof(norm0, ¬ p ∨ (q ∧ r) ...
  fof(norm1, (¬ p ∨ q) ∧ (¬ p ∨ r), inf(clausify, [norm0])).
\end{verbatim}

>>>>

\subsubsection{Clausify.}
\label{sssec:clausification}

The \clausify rule is another alternative way to transform a formula
into its clausal normal form but without any simplification since
remember this conversion is taking into account by the \canonicalize
rule. It is important to notice that this kind of conversion between
one formula to its clausal normal form is not unique, and \Metis has
a customized approach to perform that transformation. Hence, we
perform a reordering of the conjunctive normal form for the source
formula.

\begin{myexamplenum}
In the following \TSTP derivation by \Metis, we see how
\clausify transforms the \texttt{n₀} formula to get \texttt{n₁} formula.

\begin{verbatim}
  fof(n₀, ¬ p ∨ (q ∧ r) ...
  fof(n₁, (¬ p ∨ q) ∧ (¬ p ∨ r), inf(clausify, n₀)).
\end{verbatim}

### 2017-11-20 paper/sections/proof-reconstruction/resolve.tex
[ paper ] lots of changes.

\subsubsection{Resolution.}
\label{sssec:resolve}

@@

\begin{equation}
\label{eq:fassoc}
  \begin{aligned}
    &\hspace{.495mm}\fassoc_{⬓} : \Prop \to \Prop\\
    &\begin{array}{lll}
    \fassoc_{⬓} &((φ₁~⬓~φ₂)~⬓~φ₃) &= \fassoc_{⬓}~(φ₁~⬓~(φ₂~⬓~φ₃))\\
    \fassoc_{⬓} &(φ₁~⬓~φ₂)        &= φ₁~∨~\fassoc_{⬓}~φ₂\\
    \fassoc_{⬓} &φ                &= φ.
    \end{array}
  \end{aligned}
\end{equation}

@@

\begin{myremark}
In \TPTP syntax, the formulas are in left-associative form by default.
Despite of that convention, \Metis assumes the formulas to be in
right-associative form by default. This is a matter to take into account for
the proof-reconstruction.
\end{myremark}

@@

The $\ffactor$ function in~\eqref{eq:factor-definition}
simplifies a special case of disjunction, the repeated disjuncts
(\eg, $\ffactor~(φ ∨ φ)~=~φ$).
Notice that other cases like $φ~∨~(ψ~∨~φ)$ do no reduce to $ψ~∨~φ$.
We use this function in Lemma~\ref{lem:sbuild-or}.

@@

\begin{myremark}
Notice that using $\fsbuild_{∨}$ we can build not only a disjunction
with the same disjuncts of the source formula but also a complete different
disjunction by adding new disjuncts to the source formula
via introduction rules for disjunctions.
\end{myremark}

@@

\begin{myexamplenum}\hspace{10cm}
\label{ex:resolve-tstp}
\begin{verbatim}
  cnf(ref4, (¬ r ∨ p ∨ q), inf(canonicalize, [norm5])).
  cnf(ref5, (p ∨ q ∨ r), inf(canonicalize, [norm6])).
  cnf(ref6, (p ∨ q), inf(resolve, r, [ref5, ref4])).
\end{verbatim}
In the excerpt above, we apply resolution to the first two formulas,
$¬ r ∨ p ∨ q$ and $p ∨ q ∨ r$. The last line tells us the literal used
for resolution is $r$. Syntactically speaking,
we can not derive neither the conclusion $p ∨ q$ in \verb!ref6! nor apply the resolution theorem with \verb!ref4! and \verb!ref5! since the formulas do not fit the pattern required.

If the scenario would have other like replacing \verb!ref5! by
\begin{verbatim}
  cnf(ref5, (r ∨ p ∨ q), inf(canonicalize, [norm6])).
\end{verbatim}
The \resolve rule have could derive $(p ∨ q) ∨ (p ∨ q)$, but again, that is not the expected result.
\end{myexamplenum}

@@

\begin{myexamplenum}
Continuing with the problem presented in Example~\ref{ex:resolve-tstp},
we can use Theorem \ref{thm:resolve} to derive $Γ ⊢ p ∨ q$.
% Let be $φ₁ ≡ p ∨ q ∨ r$ and $φ₂ ≡ ¬ r ∨ p ∨ q$.

>>>>

\subsubsection{Resolve.}
\label{sssec:resolve}

@@

\begin{equation}
\label{eq:fassoc}
  \begin{aligned}
    &\hspace{.495mm}\fassoc_{⬓} : \Prop \to \Prop\\
    &\begin{array}{lll}
    \fassoc_{⬓} &((φ₁~⬓~φ₂)~⬓~φ₃) &= \fassoc_{⬓}~(φ₁~⬓~(φ₂~⬓~φ₃))\\
    \fassoc_{⬓} &(φ₁~⬓~φ₂)        &= φ₁~⬓~\fassoc_{⬓}~φ₂\\
    \fassoc_{⬓} &φ                &= φ.
    \end{array}
  \end{aligned}
\end{equation}

@@

\begin{myremark}

In \TPTP syntax, the formulas are in left-associative form by
default. Despite of that convention, \Metis assumes the formulas to
be in right-associative form by default. This is a matter to take
into account for the proof-reconstruction.

\end{myremark}

@@

The $\ffactor$ function in~\eqref{eq:factor-definition}
simplifies a special case of disjunction, the repeated disjuncts
(\eg, $\ffactor~(φ ∨ φ)~=~φ$).
Notice that other cases like $φ~∨~(ψ~∨~φ)$ do not reduce to $ψ~∨~φ$.
We use this function in Lemma~\ref{lem:sbuild-or}.

@@

\begin{myremark}

Notice that using $\fsbuild_{∨}$ we can build not only a disjunction
with the same disjuncts of the source formula but also a complete
different disjunction by adding new disjuncts to the source formula
via introduction rules for disjunctions.

\end{myremark}

@@

\begin{myexamplenum}\hspace{10cm}
\label{ex:resolve-tstp}
\begin{verbatim}
  cnf(r₄, (¬ r ∨ p ∨ q), ...
  cnf(r₅, (p ∨ q ∨ r), ...
  cnf(r₆, (p ∨ q), inf(resolve, r, [r₅, r₄])).
\end{verbatim}

In the excerpt above, we apply resolution to the first two formulas,
$¬ r ∨ p ∨ q$ and $p ∨ q ∨ r$.
The last line tells us the literal used
for resolution is $r$. Syntactically speaking, we can not derive
neither the conclusion $p ∨ q$ in \verb!r₆! nor apply the resolution
theorem with \verb!r₄! and \verb!r₅! since the formulas do not fit
the pattern required.

If the scenario would have other like replacing \verb!r₅! by
\begin{verbatim}
  cnf(r₅, (r ∨ p ∨ q), ...
\end{verbatim}
The \resolve rule have could derive $(p ∨ q) ∨ (p ∨ q)$, but again, that is not the expected result.
\end{myexamplenum}

@@

\begin{myexamplenum}

Continuing with the problem presented in
Example~\ref{ex:resolve-tstp}, we can use Theorem \ref{thm:resolve} to derive $Γ ⊢ p ∨ q$.

### 2017-11-20 paper/sections/proof-reconstruction/strip.tex
[ paper ] lots of changes.

\subsubsection{Stripping a goal.}
\label{sssec:strip-a-goal}

@@

\begin{verbatim}
  fof(goal, conjecture, (p ∧ r) ∧ q)).
  fof(sg0, p, inf(strip, [goal])).
  fof(sg1, p ⇒ r, inf(strip, [goal])).
  fof(sg2, (p ∧ r) ⇒ q, inf(strip, [goal])).
\end{verbatim}

@@

In a similar way as we define $\fuh_{cm}$ in~\eqref{eq:uh-complexity}, we
define $\fstrip_{cm}$ as the complexity measure for the $\fstrip₀$ function.
Then we define the \fstrip function as follows in~\ref{eq:strip}.

@@

\begin{mainth}
\label{thm:strip}
Let $\text{n} : \Nat$ be the complexity measure of the strip function defined
in~\eqref{eq:strip-fixed}.
Let $s₁$ and $s₂$ be the subgoals of the goal $φ$, that is,
$$\fstrip₁~φ~n~≡~s₁∧~s₂.$$
If $Γ ⊢ s₁$ and $Γ ⊢ s₂$ then $Γ ⊢ φ$.
\end{mainth}

@@

\begin{myexamplenum}
In the following \TSTP derivation, note that the three subgoals
\verb!sg0!, \verb!sg1! and \verb!sg2! are assumed by the \negate
rule in \verb!neg0!, \verb!neg1! and  \verb!neg2! respectively.

\begin{verbatim}
  fof(goal, conjecture, (p ∧ r ∧ q)).
  fof(sg0, p, inf(strip,[goal])).
  fof(sg1, p ⇒ r, inf(strip,[goal])).
  fof(sg2, (p ∧ r) ⇒ q, inf(strip,[goal])).
  fof(neg0, ¬ p, inf(negate,[sg0])).
  ...
  fof(neg1, ¬ (p ⇒ r), inf(negate,[sg1])).
  ...
  fof(neg2, ¬ ((p ∧ r) ⇒ q), inf(negate,[sg2])).
\end{verbatim}
\end{myexamplenum}

>>>>

\subsubsection{Strip.}
\label{sssec:strip-a-goal}

@@

\begin{verbatim}
  fof(goal, conjecture, (p ∧ r) ∧ q)).
  fof(s₁, p, inf(strip, goal)).
  fof(s₂, p ⇒ r, inf(strip, goal)).
  fof(s₃, (p ∧ r) ⇒ q, inf(strip, goal)).
\end{verbatim}

@@

In a similar way as we define $\fuh_{cm}$
in~\eqref{eq:uh-complexity}, we
define the $\fstrip_{cm}$ function in Appendix~\ref{app:strip-cm} as the complexity measure for the $\fstrip₀$ function.
Then we define the \fstrip function as follows in~\ref{eq:strip}.

@@

\begin{mainth}
\label{thm:strip}
Let $\text{n} : \Nat$ be the complexity measure of the strip function defined
in~\eqref{eq:strip-fixed}.
Let $s₂$ and $s₃$ be the subgoals of the goal $φ$, that is,
$$\fstrip₁~φ~n~≡~s₂∧~s₃.$$
If $Γ ⊢ s₂$ and $Γ ⊢ s₃$ then $Γ ⊢ φ$.
\end{mainth}

@@

\begin{myexamplenum}
In the following \TSTP derivation, note that the three subgoals
\verb!s₁!, \verb!s₂! and \verb!s₃! are assumed by the \negate
rule in \verb!neg₁!, \verb!neg₂! and  \verb!neg₃! respectively.

\begin{verbatim}
  fof(goal, conjecture, (p ∧ r ∧ q)).
  fof(s₁, p, inf(strip, goal)).
  fof(s₂, p ⇒ r, inf(strip, goal)).
  fof(s₃, (p ∧ r) ⇒ q, inf(strip, goal)).
  fof(neg₁, ¬ p, inf(negate, s₁)).
  ...
  fof(neg₂, ¬ (p ⇒ r), inf(negate, s₂)).
  ...
  fof(neg₃, ¬ ((p ∧ r) ⇒ q), inf(negate, s₃)).
\end{verbatim}
\end{myexamplenum}
